### 2021-04-26 tex/sections/blockchain.tex

%! Author = aybehrouz

\section{Persistence}\label{sec:persistence}
\input{sections/persistence}

\section{Transactions}\label{sec:transactions}

Argennon has four types of transaction:

\begin{itemize}
    \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
    AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
    assets, including ARGs, is done by these transactions.
    \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
    contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
    on.
    \item \texttt{unInstallApp} removes an AVM smart contract.
    \item \texttt{updateApp} updates an AVM smart contract.
\end{itemize}

All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
fee sink accounts.

Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
for more details.

\section{Blockchain}\label{sec:blockchain}

Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
if a signature of the same user with a bigger or equal nonce is already used
(i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.

To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
have the problem of "un-removable empty accounts" like Ethereum.}

The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
blockchain. Every block of the Argennon blockchain contains the following information:

\begin{center}
    \begin{tabular}{||c||}
        \hline
        Block \\ [0.5ex]
        \hline\hline
        commitment to the ZK-EDB storing heap pages \\ [0.7ex]
        commitment to the ZK-EDB storing code areas \\ [0.7ex]
        commitment to the set of transactions       \\ [0.7ex]
        previous block hash                         \\ [0.7ex]
        random seed                                 \\ [0.7ex]
        \hline
    \end{tabular}
\end{center}

For confirming a new block, nodes that are not validators only need to verify the block certificate. For
verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
emulate the AVM execution.

On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
commitments based on the modified pages and verify the commitments included in the new block. Validators also
calculate and verify the commitment to the new block's transaction set.

\note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
block, and they will update their database by emulating the AVM execution.}

\section{Consensus}\label{sec:consensus}
\input{sections/consensus}

\section{Incentive mechanism}\label{sec:incentive-mechanism}

\subsection{Transaction Fee}\label{subsec:transaction-fee}

Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
transaction fee between these two fee sink accounts.

When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
puts too many transactions in his block and the validation of the block becomes too difficult, some validators
may not be able to validate all transactions on time. If a validator can not validate a block in the required
time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
proposer is incentivized to use network transaction capacity optimally.

On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
bytes be lower than a certain threshold.}

Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
validators can not finish the execution of these transactions, the network will not be able to charge the
attacker any fees, and he would be able to waste validators resources for free.

To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
will include memory, network and processor related resources. Also, the protocol defines an execution cost for
every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
considered failed and the network can receive the proposed fee of that transaction.

Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
resources it needs:

\begin{itemize}
    \item Execution cost
    \item A list of heap/code-area locations for reading
    \item A list of heap locations for writing
    \item A list of heap pages it will deallocate
    \item Number and size of heap pages it will allocate
\end{itemize}

If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
location that is not included in its reading list, it will be considered failed and the network can receive the
proposed fee of that transaction.

\note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
used in the final emulation.}

\subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}

The incentive mechanism for ZK-EDB servers should have the following properties:

\begin{itemize}
    \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
    which are used more frequently.
    \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
    \item Making more accounts will not provide any advantages for a ZK-EDB server.
\end{itemize}

For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:

\begin{itemize}
    \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
    \item \texttt{receivedBlockID}: the ID of the received memory block.
    \item \texttt{round}: the current round number.
\end{itemize}

\note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}

To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
\texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
between all \emph{winning tickets} of the lottery.

\note{One ZK-EDB server could own multiple winning tickets in a round.}

To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:

\begin{itemize}
    \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
    \item The data block in the receipt was needed for validating the \textbf{previous} block.
    \item The receipt round number is \(r - 1\).
    \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
\end{itemize}
For selecting the winning receipts we could use a random generator:
\begin{verbatim}
IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
    the receipt issued by validatorPK for receivedBlockID is a winner
\end{verbatim}
\begin{itemize}
    \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
    \item \texttt{validatorPK} is the public key of the signer of the receipt.
    \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
    \item \texttt{winProbability} is the probability of winning in every round.
    \item \texttt{seed} is the current block seed.
    \item \texttt{|} is a concatenation operator.
\end{itemize}

\note{The winners of the lottery were validators one round before the lottery round.}

Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
challenge. Solving a round challenge requires the content of the memory block which was selected as the round
challenge. This will encourage ZK-EDBs to store all memory blocks.

A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}

The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
\(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.

\subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}

Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
address of the smart contract.

To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
\texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
\texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
the protocol considers the total size of all the memory pages that were accessed for either reading or writing
during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
AVM memory will be a linear function of \texttt{dbFee/memTraffic}

When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
This will prevent smart contracts from profit taking by trading memory with the protocol.

\section{Concurrency}\label{sec:concurrency}
\input{sections/concurrency}

\section{Smart Contract Oracle}\label{sec:smart-contract-oracle}

A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
\texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
and so on.