### 2021-06-20 tex/sections/draft.tex

%! Author = aybehrouz

In Argennon \emph{light} nodes do not keep a full copy of the AVM storage and for emulating the Argennon Virtual Machine
( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve the required pages.
Because light node need to be able to validate block certificates, they usually cache a large part of the staking
database, and keep their cache updated to make sure they can keep themselves in sync with the blockchain.

On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
execution of the Argennon Virtual Machine. To do so, first they retrieve all storage pages they
need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
transactions included in the new block. This will modify some pages of the AVM storage, so they update the ZK-EDB
commitments based on the modified pages and verify the commitments included in the new block. Validators also
calculate and verify the commitment to the new block's transaction set.

\note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
block, and they will update their database by emulating the AVM execution.}

The only information that all Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
blockchain.
We store the commitment of this
transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
if a signature of the same user with a bigger or equal nonce is already used
(i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.

To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
have the problem of "un-removable empty accounts" like Ethereum.}

In our system a newly created account will not have voting power for some time, no matter how high its
balance is. While this is a desirable property, in case a large proportion of total system tokens are
transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
the degree of decentralization in our system.

However, this situation is easily detectable by comparing the total stake of the system with the total balance of
users. If after confirming a block the total stake of the system goes too low and we have:
\[
    \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
\]
The protocol will perform a \emph{time shift} in the system: the time step of the system
will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
for new accounts with a non-zero balance, giving them more influence in the agreement protocol.

For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
\(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
\[
    S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
\]
Therefore, we have:
\[
    \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
\]
This equation shows that when the balance of users is not changing over time the total stake of the system is the
exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
calculate the new total stake of the system from the following equation:

\[
    \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
\]
Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
\[
    m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
\]

An Argennon block certificate is not a part of the block data. This means that the Argennon blockchain does not have a
unique certificate sequence. Any node may have its own valid certificate sequence based on the signatures
it has collected over time.

\subsection{Memory Shards}\label{subsec:memory-shards}

In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{shards}.
Each memory shard can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
consensus on new values of the commitment of any shard can be achieved by a different voting committee.

If a transaction does not modify a memory shard and in the transaction ordering of the block it comes after
any transaction which modifies that shard, then the execution of that transaction is not needed for calculating
the new commitment of the shard. Consequently, the voting committee of that memory shard can safely ignore such a
transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
we see in Algorithm~\ref{alg:prune_dag}.

%##\includegraphics[width=17cm]{../img/Alg4s.png}
\begin{algorithm}
    \DontPrintSemicolon
    \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Shard}{$Sh_k$}\SetKwData{Txns}{$T$}
    \SetKwFunction{RPrune}{RecursivePrune}
    \SetKwProg{Fn}{Function}{}{}
    \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
    \Input{An execution dag \Graph and a memory shard \Shard}
    \Output{The pruned execution dag based on \Shard}
    \BlankLine
    \For{each vertex \V with out degree $0$}
    {
        \RPrune{\V}\;
    }
    \BlankLine
    \Fn{\RPrune{\V}}
    {
        \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Shard}
        {
            remove \V from \Graph\;
            \For{each $u$ such that edge $(u,\V)$ was in \Graph}
            {
                \RPrune{u}\;
            }
        }
    }
    \caption{Pruning an execution DAG}\label{alg:prune_dag}
\end{algorithm}

If we choose shards in a way that most transactions only modify memory locations of one shard,
likely many transactions of a block only need to be validated by one voting committee and can be validated in
parallel by different committees.

Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
that having multiple voting committees will not change the security properties of the Argennon agreement protocol.