### 2022-07-06 tex/A.tex
Update abstract
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

%! suppress = TooLargeSection
%! Author = aybehrouz
%! Date = 2/28/21

\documentclass[11pt, a4paper]{report}
\usepackage[a4paper]{geometry}

% Packages
\usepackage{tipa}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage[ruled,boxed]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{tcolorbox}
\usepackage{amsmath}

%\binoppenalty=99999
%\relpenalty=99999
\newcommand{\note}[1] {
    \begin{tcolorbox}[colframe=white,colback=white]
        \emph{#1}
        %\emph{#1}\medskip
    \end{tcolorbox}
}

\lstset{
    language=Java,
    morekeywords={main, initialize, constructor, external, virtual, Map, Account, signature, overrides, is, string},
    basicstyle=\small\sffamily,
    columns=fullflexible,
%   keywordstyle=\bfseries,
    breaklines=true,
    showstringspaces=false,
    mathescape
}

\title{Argennon: A Scalable Cloud Based Smart Contract Platform}
\author{aybehrouz}
\date{January 2021}

\begin{document}
    \maketitle
    \begin{abstract}
        Argennon is a next generation cloud based blockchain and smart
        contract platform. The Argennon blockchain uses
        a hybrid proof of stake (HPoS) consensus protocol, which is capable of combining the benefits of
        a centralized and a decentralized system. Thanks to the cloud
        based design of the Argennon blockchain, transaction validation does not require costly
        computational resources, and normal personal computers or smartphones, with limited hardware capabilities, can
        validate transactions and actively
        participate in the Argennon consensus protocol. This property makes Argennon a truly decentralized and
        democratic blockchain and one of the most secure existing platforms.

        The Argennon cloud is formed by a permission-less network of Publicly Verifiable Database (PV-DB) servers. A
        PV-DB server in Argennon, is a conventional data server which uses its computational and
        storage resources to help the Argennon network process transactions. A large portion of incentive rewards in
        the Argennon platform is devoted to PV-DB servers. This incentivizes the development of conventional
        networking, storage and computational hardware, which can benefit all areas of information technology.

        By using cryptographic accumulators, the integrity of data on the Argennon cloud is guaranteed, and there
        is no need for trusting cloud servers. At the same time, by using a smart clustering algorithm the network usage
        and the overhead of the cryptographic commitment scheme is kept manageable.
    \end{abstract}
    \tableofcontents

    \chapter{The Argennon Smart Contract Execution Environment}\label{ch:AVM}
    \input{sections/avm}

    \chapter{Persistence Layer}\label{ch:persistance}
    \input{sections/persistence}

    \chapter{Networking Layer}\label{ch:networking}

    \section{Normal Mode}\label{sec:normal-mode}
    Unlike conventional blockchains, Argennon does not use a P2P network architecture. Instead, it uses a
    client-server topology, based on a permission-less list of ZK-EDB servers. ZK-EDB servers are a
    crucial part of the Argennon ecosystem, and they form the backbone of the Argennon networking layer.
    \note{not yet written...}

    \section{Censorship Resilient Mode}\label{sec:cens-res-mode}
    \note{not yet written...}

    \chapter{The Argennon Blockchain}\label{ch:argennon-blockchain}

    \section{Blocks}\label{sec:blocks}
    \input{sections/blocks}

    \section{Consensus}\label{sec:consensus}
    \input{sections/consensus}

    \input{sections/elements}

    \section{Incentive mechanism}\label{sec:incentive-mechanism}
    \input{sections/incentive}

    \chapter{Governance}\label{ch:governance}

    \section{ADAGs}\label{sec:adags}
    The Argennon Decentralized Autonomous Governance system (ADAGs)
    \note{not yet written...}

    \chapter{The Argon Language}\label{ch:argon-lang}
    \input{sections/argon}
\end{document}