### 2022-08-07 tex/A.tex
changed chapter 1 of whitepaper

%! suppress = TooLargeSection
%! Author = aybehrouz
%! Date = 2/28/21

\documentclass[11pt, a4paper]{report}
\usepackage[a4paper]{geometry}

% Packages
\usepackage{tipa}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage[ruled,boxed]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{tcolorbox}
\usepackage{amsmath}
\usepackage{mathtools}

%\binoppenalty=99999
%\relpenalty=99999
\newcommand{\note}[1] {
    \begin{tcolorbox}[colframe=white,colback=white]
        \emph{#1}
        %\emph{#1}\medskip
    \end{tcolorbox}
}

\lstset{
    language=Java,
    morekeywords={main, initialize, constructor, external, virtual, Map, Account, signature, overrides, is, string},
    basicstyle=\small\sffamily,
    columns=fullflexible,
%   keywordstyle=\bfseries,
    breaklines=true,
    showstringspaces=false,
    mathescape
}

\title{Argennon: A Scalable Cloud Based Smart Contract Platform}
\author{aybehrouz}
\date{January 2021}

\begin{document}
    \maketitle
    \begin{abstract}
        Argennon is a next generation cloud based blockchain and smart
        contract platform. The Argennon blockchain uses
        a hybrid proof of stake (HPoS) consensus protocol, which is capable of combining the benefits of
        a centralized and a decentralized system. In Argennon, ledger storage and transaction processing are
        outsourced to the Argennon cloud and normal personal computers or smartphones, with limited hardware
        capabilities, can validate transactions and actively
        participate in the consensus protocol. This property makes Argennon a truly decentralized and
        democratic blockchain and one of the most secure existing platforms.

        The Argennon cloud consists of a Committee of Delegates (CoD) and Publicly Verifiable Cloud (PVC) servers. 
        The CoD is elected per term by Proof of Stake and is responsible for processing transactions and creating blocks.
        After validator assemblies have verified the validity of these blocks they are stored in the PVC servers.
        PVC servers are publicly verifiable. Computational Integrity (CI) is
        achieved by using succinct cryptographic proofs (STARK/SNARK)
        and data integrity is guaranteed by cryptographic accumulators. At the same time, a smart
        clustering algorithm keeps the bandwidth usage and the overhead of cryptography manageable
        for validators.

        This set up allows for very high transaction throughput while at the same time remaining highly secure. The network 
        stays safe as long as at least one delegate and at least 1/3 of the validators in all validator assemblies remain honest.
    \end{abstract}
    \tableofcontents

    \chapter{Introduction}\label{ch:intro}
    \input{sections/intro}

    \chapter{The Argennon Smart Contract Execution Environment}\label{ch:AVM}
    \input{sections/avm}

    \chapter{Persistence Layer}\label{ch:persistance}
    \input{sections/persistence}

    \chapter{Networking Layer}\label{ch:networking}

    \section{Normal Mode}\label{sec:normal-mode}
    Unlike conventional blockchains, Argennon does not use a P2P network architecture. Instead, it uses a
    client-server topology, based on a permission-less list of ZK-EDB servers. ZK-EDB servers are a
    crucial part of the Argennon ecosystem, and they form the backbone of the Argennon networking layer.
    \note{not yet written...}

    \section{Censorship Resilient Mode}\label{sec:cens-res-mode}
    \note{not yet written...}

    \chapter{The Argennon Blockchain}\label{ch:argennon-blockchain}

    \section{Blocks}\label{sec:blocks}
    \input{sections/blocks}

    \section{Consensus}\label{sec:consensus}
    \input{sections/consensus}

    \input{sections/elements}

    \section{Incentive mechanism}\label{sec:incentive-mechanism}
    \input{sections/incentive}

    \chapter{Governance}\label{ch:governance}

    \section{ADAGs}\label{sec:adags}
    The Argennon Decentralized Autonomous Governance system (ADAGs)
    \note{not yet written...}

    \chapter{The Argon Language}\label{ch:argon-lang}
    \input{sections/argon}
\end{document}